---
title: "Regression Analysis"
author: "Jason Reese"
---

##Getting Data
```{r library, include=FALSE}
library(pander)
library(lessR)
```

```{r read/subset data, include=FALSE}
pass <- Read("https://raw.githubusercontent.com/Jsr5126/DS-Portfolio/0f46a9ea57eb9298a07604bc176757bd9c6cdb5b/2020Passing.txt")
pass_s <- pass[c(1, 2, 3, 10, 15, 19, 25)]
pass_s <- pass_s[1:50,]
```
The purpose of this project is doing a regression analysis of NFL passing stats from the 2020 season. The three predictor variables  used to do the analysis were volume (Attempts), efficiency (Yards per Attempt), and competition (Sacks). Also, the three values roughly represent the system the quarterback is in (Attempts), their own abilities (Yards per Attempt), and strength of schedule (Sacks). Then, I selected the top 50 players in the league by attempts. This subset was successful in eliminating non-quarterbacks and, for the most part, eliminated players that did not start the game.

##Multiple Regression
```{r head}
pander(head(pass_s))
```
```{r linear regression model}
fit <- lm(Int ~ Att + Y.A + Sk, pass_s)
pander(fit)
```
With a p-value less than 0.05 in all three predictors I can reject the null hypothesis. My equation is:
\begin{align}Interceptions = 9.777 + Att(0.01438) + Y.A(-1.27) + Sk(0.077710)\end{align}
This means that increasing the number of attempts or the number of sacks or decreasing the yards per attempt will increase the number of interceptions. In determining interceptions, sacks are the least significant and attempts are the most significant. The upper and lower confidence intervals here are misleading because they contain players that started as few as 0 or 1 games. R2 being the measure of how close the data is fitted to the regression. A value of 0.6494 multiplied by the sample (50 players) would result in 32.47. This closely represents the number of starting caliber quarterbacks in the league (32) undoubtedly because NFL managers are aware of analysis like this and would drop their QB if they fall below expected metrics of the quarterbacks available to start.

##Plot
```{r plot models}
 plot(fit)
```
